Synthesis report

### 2020-08-05 analysis/reports/less_than_equal.Rmd
move old reporst to old_reports directory

# Illustrations of 95% interval

To show the 95% interval, we need to load the distribution of shape metric values from the samples from the feasible set for a few communities.

# Skewness and evenness histograms by dataset

```{r first hists}

ggplot(filter(all_di, s0 > 2), aes(skew_percentile)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 95, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Skewness, all sites")

ggplot(filter(all_di, s0 > 2, log_nparts > log(50)), aes(skew_percentile)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 95, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Skewness, all sites")

ggplot(all_di, aes(simpson_percentile)) + 
 geom_histogram(bins = 100) +
  geom_vline(xintercept = 5, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Evenness, all sites")
ggplot(filter(all_di, log_nparts > log(50)), aes(simpson_percentile)) + 
 geom_histogram(bins = 100) +
  geom_vline(xintercept = 5, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Evenness, all sites")
```

# Proportion of sites with highly skewed or uneven SADs

```{r results table}

overall_skew_results <- all_di %>%
  filter(s0 > 2) %>%
  mutate(high_skew = skew_percentile > 95) %>%
  group_by(dat) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results <- all_di %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(dat) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

overall_results <- left_join(overall_skew_results, overall_simpson_results)

overall_results

overall_skew_results2 <- all_di %>%
  filter(s0 > 2, log_nparts > log(50)) %>%
  mutate(high_skew = skew_percentile > 95) %>%
  group_by(dat) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results2 <- all_di %>%
  filter(log_nparts > log(50)) %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(dat) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

overall_results2 <- left_join(overall_skew_results2, overall_simpson_results2)

overall_results2

@@

There are fewer sites for skewness because we exclude sites with only 2 species.

# Size of feasible set for each dataset

```{r fs size hists}

all_di <- all_di %>%
  mutate(log_nparts_truncated = ifelse(
    log_nparts > 50,
    50,
    log_nparts)
  )

ggplot(all_di, aes(log_nparts_truncated)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y") +
  theme_bw()

fs_size_results <- all_di %>%
  group_by(dat) %>%
  summarize(prop_smaller_than_5=  mean(log_nparts <= 5),
            prop_smaller_than_7 =  mean(log_nparts <= 7),
            prop_smaller_than_10 = mean(log_nparts <= 10))
fs_size_results

@@

FIA is the only dataset that is overwhelmingly dominated by sites with fewer than `r exp(10)` elements in the feasible set. MCDB also has many small sites, but these are counterbalanced by larger ones.

62% of the FIA sites have fewer than `r exp(7)` elements in the feasible set, and 30% have fewer than `r exp(5)`.

```{r fs size to s0 and n0, fig.dim = c(3,3)}
ggplot(all_di, aes(x = log_s0, y = log_n0, color = log_nparts <= 5)) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "top")

ggplot(all_di, aes(x = log_s0, y = log_n0, color = log_nparts <= 7)) +
  geom_point() +
  theme_bw()+
  theme(legend.position = "top")

ggplot(all_di, aes(x = log_s0, y = log_n0, color = log_nparts <= 10)) +
  geom_point() +
  theme_bw()+
  theme(legend.position = "top")
```

Having fewer than 150 elements ~ fewer than 3 species, or fewer than `r exp(2.75)` individuals  with up to 20 species.

Fewer than 1000 elements ~ fewer than 3 species, or fewer than `r exp(3.25)` individuals with up to 20 species.

Fewer than 20000 elements ~ fewer than 4 species, or fewer than `r exp(4)` individuals.

# 95% interval for small versus large communities

```{r 95 interval}

ggplot(all_di, aes(x = log_nparts, y = skew_95_ratio_1t)) +
  geom_point(alpha = .2) + theme_bw() + ggtitle("Skewness 95% interval")

ggplot(all_di, aes(x = log_nparts, y = skew_95_ratio_1t)) +
  geom_point(alpha = .01) +
  xlim(0, 15) +
  theme_bw() +
  ggtitle("Skewness 95% interval - only small feasible sets")

ggplot(all_di, aes(x = log_nparts, y = simpson_95_ratio_1t)) +
  geom_point(alpha = .2) + theme_bw()+ ggtitle("Evenness 95% interval")

ggplot(all_di, aes(x = log_nparts, y = simpson_95_ratio_1t)) +
  geom_point(alpha = .05) +
  xlim(0, 15) +
  theme_bw()+
  ggtitle("Evenness 95% interval - only small feasible sets")

The 95% ratio decreases for large FS.

For skewness, it is still quite high even up to 20,000 elements in the FS. But it is starting to decline by then?

# Comparison of FIA to similarly-sized communities

```{r fia sized}
fia_max_s <- max(filter(all_di, dat == "fia")$s0)
fia_max_n <- max(filter(all_di, dat == "fia")$n0)

fia_max_nparts <-max(filter(all_di, dat == "fia")$log_nparts)

fia_sized <- filter(all_di, s0 <= fia_max_s, n0 <= fia_max_n) %>%
  mutate(fia_yn = dat == "fia") %>%
  mutate(fia_yn = ifelse(fia_yn, "FIA", "Other datasets")) %>%
  mutate(right_corner = FALSE)

ggplot(filter(fia_sized, s0 > 2, nparts > 50), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y")

ggplot(fia_sized, aes(x = log_nparts)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  geom_vline(xintercept = log(50))

ggplot(fia_sized, aes(s0, n0, color = fia_yn)) +
  geom_point()

# Removing the smallest communities from FIA

```{r fia no small}

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts < exp(7)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("50-1,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts >= exp(7), nparts < exp(10)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("1,000 - 20,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts >= exp(10)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("More than 20,000 parts in FS")
```

>>>>

```{r skew change}

ggplot(filter(all_di, s0 > 2), aes(x = skew_percentile_excl, y = skew_percentile, color = nsamples)) +
  geom_point() +
  facet_wrap(vars(dat))

all_di <- all_di  %>%
  mutate(skew_change = skew_percentile - skew_percentile_excl,
         even_change = simpson_percentile -
           simpson_percentile_excl)

ggplot(filter(all_di, s0 > 2), aes(x = log_nparts, y = skew_change)) +
  geom_point() +
  xlim(0,15) +
  facet_wrap(vars(dat))

@@

```{r hists}

@@

### 2020-08-05 analysis/reports/manuscript_main.Rmd
move old reporst to old_reports directory

ggplot(fia_sized, aes(x = log_nparts)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  geom_vline(xintercept = log(50))

ggplot(fia_sized, aes(s0, n0, color = fia_yn)) +
  geom_point()

@@

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts < exp(7)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("50-1,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts >= exp(7)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("More than 1,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts >= exp(10)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("More than 20,000 parts in FS")
```

>>>>

ggplot(filter(fia_sized, s0 > 2, nparts > 50), aes(skew_percentile_excl)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y")

ggplot(filter(fia_sized, s0 > 2), aes(skew_percentile_excl)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y")
ggplot(fia_sized, aes(x = log_nparts)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  geom_vline(xintercept = log(50))

fia_sized %>%
  filter(s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile > 95),
            nsites = dplyr::n())

fia_sized %>%
  filter(s0 > 2, nparts > 50) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile > 95),
            nsites = dplyr::n())

fia_sized %>%
  group_by(fia_yn) %>%
  summarize(prop_even_low = mean(simpson_percentile < 5),
            nsites = dplyr::n())

fia_sized %>%
    filter(s0 > 2, nparts > 50) %>%
  group_by(fia_yn) %>%
  summarize(prop_even_low = mean(simpson_percentile < 5),
            nsites = dplyr::n())

@@

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts < exp(7)), aes(skew_percentile)) +
  geom_histogram(bins = 50) + theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("50-1,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts < exp(7)), aes(x = log(s0), y = log(n0), color = fia_yn)) +
  geom_point(alpha = .2) +
  facet_wrap(vars(fia_yn))

ggplot(filter(fia_sized, s0 > 2, nparts >= exp(7), nparts < exp(10)), aes(skew_percentile)) +
  geom_histogram(bins = 50) + theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("1,000 - 20,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts >= exp(7), nparts < exp(10)), aes(x = log(s0), y = log(n0), color = fia_yn)) +
  geom_point(alpha = .2) +
  facet_wrap(vars(fia_yn))

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts >= exp(10)), aes(skew_percentile)) +
  geom_histogram(bins = 50) + theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("More than 20,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts >= exp(7), nparts < exp(10)), aes(x = log(s0), y = log(n0), color = fia_yn)) +
  geom_point(alpha = .2) +
  facet_wrap(vars(fia_yn))
```

### 2020-08-05 analysis/reports/old_reports/fia_size.Rmd
move old reporst to old_reports directory

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
all_di_old <- read.csv(here::here("all_di.csv"), stringsAsFactors = F)

@@

fia_sized <- filter(all_di, s0 <= fia_max_s, !singletons, n0 <= fia_max_n) %>%
  mutate(fia_yn = dat == "fia") %>%
  mutate(random_5th = as.integer(sample(2, size = n(), replace = T, prob = c(0.035, 0.965))))

>>>>

all_di <- read.csv(here::here("analysis", "reports", "all_di_old.csv"), stringsAsFactors = F)
all_di_old <- read.csv(here::here("all_di.csv"), stringsAsFactors = F)

@@

fia_sized <- filter(all_di, s0 <= fia_max_s, !singletons, n0 <= fia_max_n, s0 > 2) %>%
  mutate(fia_yn = dat == "fia") %>%
  mutate(random_5th = as.integer(sample(2, size = n(), replace = T, prob = c(0.035, 0.965))))
